### 2015-10-16 figures/example/example-body.tex
revise figures
change the units of x axis… (strange setting of my student)

		\renewcommand{\arraystretch}{1.3}

		\def\ux{1.1cm}\def\uy{0.75cm} 
		\begin{tikzpicture}[x=\ux,y=\uy, font=\sffamily,thick]  
		\tikzset{
			task/.style={fill=#1,  rectangle, text height=.3cm},
			task1a/.style={task=green!30},
			task1b/.style={task=green},
			task2a/.style={task=orange!30},
			task2b/.style={task=orange, minimum width=1mm},
			task3a/.style={task=pink, minimum width=1mm},
			task3b/.style={task=pink!80, minimum width=1mm},
			task4a/.style={task=cyan, minimum width=1mm},
			task4b/.style={task=cyan!50},
			task5/.style={task=blue},
			task6/.style={task=purple},
			task7/.style={minimum height=\uy,draw},
			task8/.style={minimum height=\uy,draw,thick},
			task9/.style={task=gray,minimum height=0.7cm,draw},
		}
		\tikzstyle{jobs}=[ fill=black!50];

		\begin{scope}[shift={(0,0)}]

		\draw[<-](0,2) -- (0,0);
		\draw[<->](5,2) -- (5,0);
		\draw[<->](10,2) -- (10,0);

		\draw[->] (0,0) node[anchor=east] {$\tau_1$}-- coordinate (xaxis) (12.5,0);

		\node[task7,minimum width=\ux, 
                anchor=south west] at (0,0){};
		\node[task7, minimum width=\ux,
		anchor=south west]at (5, 0){};
		\node[task7, minimum width=\ux,
		anchor=south west]at (10, 0){};
		\end{scope}

		\begin{scope}[shift={(0,-2.5)}]
		%%timeline

		\draw[<-](0,2) -- (0,0);
		\draw[<->](5.5,2) -- (5.5,0);
		\draw[<->](11,2) -- (11,0);

		\node[task7, minimum width=.5*\ux,
		anchor=south west] at (1, 0){\tiny $C_2^1$};
		\draw[(-), thin] (1.5, 0.3) -- (3, 0.3)
                node[anchor=south]{\scriptsize suspension} -- (4.5, 0.3);
		\node[task7, minimum width=.5*\ux,
		anchor=south west] at (4.5, 0){\tiny $C_2^2$};

		\node[task7, minimum width=.5*\ux,
		anchor=south west] at (6, 0){\tiny $C_2^1$};
		\draw[(-), thin] (6.5, 0.3) -- (8, 0.3)
                node[anchor=south]{\scriptsize suspension} -- (9.5, 0.3);

		\draw[->] (0,0)node[anchor=east] {$\tau_2$} -- coordinate (xaxis) (12.5,0);

		\node[task7, minimum width=.5*\ux,
		anchor=south west] at (11, 0){\tiny $C_2^2$};
		\node[task7, minimum width=.5*\ux,
		anchor=south west] at (11.5, 0){\tiny $C_2^1$};

               \foreach \x in {0,...,24}{
                        \draw[-](.5*\x,0.1) -- (.5*\x,-0.1)
                        node[below] {\small $\x$};

                }	
                \draw (11,1.1) node[anchor=south] {\scriptsize\textcolor{red}{deadline miss}};
		\end{scope}

		\end{tikzpicture}