### 2015-10-16 LITES/LITES-Paper.tex
restructure for sync with Authorea

\input{../preamble-LCTES}

\input{../header.tex}

\title{A Note on the Period Enforcer Algorithm  for Self-Suspending Tasks}

\author[1]{Jian-Jia Chen}
\author[2]{Bj\"orn B.\ Brandenburg}
\affil[1]{TU Dortmund University, Germany\\
  \texttt{jian-jia.chen@cs.uni-dortmund.de}}
\affil[2]{Max Planck Institute for Software Systems (MPI-SWS)\\
  \texttt{bbb@mpi-sws.org}}
\authorrunning{Chen and Brandenburg} %mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et. al.'

\Copyright{Chen and Brandenburg}%mandatory. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

%\subjclass{MANDATORY:  Please refer to \url{www.acm.org/about/class/2012}}% mandatory: Please choose 2012 ACM classifications from http://www.acm.org/about/class/2012 . E.g., cite as "Integrated circuits". 
%\keywords{MANDATORY: Please provide 1--5 keywords as a comma-separated list}% mandatory: Please provide 1-5 keywords
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%EditorialOffice macros (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Volume{XXX}
\Issue{YYY}
\DateSubmission{Date of submission}
\DateAcceptance{Date of acceptance}
\DatePublished{Date of publishing}
\SectionAreaEditor{LITES section area editor}
\DOI{10.4230/LITES.xxx.yyy.p}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\maketitle
\input{../abstract.tex}
\input{../untitled.tex}

%\input{../figures/example-original/example.png}
%\input{../figures/example/example.png}
\input{../transformation.tex}
\input{../locking.tex}
\input{../conclusion.tex}

\bibliography{../bibliography/biblio}{}

\end{document}