### 2016-02-29 conclusion.tex
ack
add ack sfb876

\section{Concluding Remarks}
\label{sec:conclusion}

We have revisited the underlying assumptions and limitations of the period enforcer algorithm, which Rajkumar \cite{Raj:suspension1991} introduced to handle segmented self-suspending real-time tasks.

One key assumption in the original proposal \cite{Raj:suspension1991} is that a deferrable task $\tau_i$ can defer its entire execution time but not parts of it. This creates some mismatches between the original self-suspending task set and the corresponding deferrable task set, which we have demonstrated with an example that shows that Theorem 5 in \cite{Raj:suspension1991} does not reflect the schedulability of the original self-suspending task system.

Furthermore, the original proposal \cite{Raj:suspension1991} left open the question of how to convert a segmented self-suspending task set to a corresponding set of deferrable tasks. Taking into account recent developments~\cite{ecrts15nelissen}, we have observed that such a task set transformation is non-trivial in the general case.

Finally, we have demonstrated that substantial difficulties arise if one attempts to combine suspension-based locks with period enforcement. These difficulties stem from the fact that period enforcement can increase contention or lock-holding times, which increases the lengths of self-suspension intervals, which then in turn feeds back into the period enforcer's minimum suspension lengths. As a consequence, period enforcement invalidates all existing blocking analyses.

Nevertheless, Theorem 5 in \cite{Raj:suspension1991} could be useful for handling self-suspending tasks (that do not use suspension-based locks) if there exist \emph{efficient} schedulability tests for the corresponding deferrable task systems or the period enforcer algorithm. However, such tests have not been found yet and the development of a precise and efficient schedulability test for self-suspending tasks remains an open problem.

\section*{Acknowledgements}

We thank James H.\ Anderson and Raj Rajkumar for their comments on drafts of this paper.
This paper has been supported by DFG, as part of the Collaborative
Research Center SFB876 (http://sfb876.tu-dortmund.de/).